---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# IPEDS

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/IPEDS)](https://CRAN.R-project.org/package=IPEDS)
<!-- badges: end -->

`IPEDS` is a package that contains datasets from The Integrated Postsecondary Education Data System (IPEDS). IPEDS is a system of interrelated surveys conducted annually by the U.S. Department of Education's National Center for Education Statistics (NCES). IPEDS annually gathers information from about 6,400 colleges, universities, and technical and vocational institutions that participate in the federal student aid programs.

## Installation

You can install the development version of IPEDS from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ajhaller/IPEDS")
```

## Load

```{r}
library(IPEDS)
```

## Available Tables

-   `adm2020`: dataset of Admissions and Test Scores for Fall 2020
-   `admin2020`: dataset of Administration for 2020
-   `complete2020`: dataset of Completions in 2020
-   `dir_info2020`: dataset of Directory Information for 2020
-   `fall_enroll2020`: dataset of Fall Enrollment for 2020
-   `fin_aid1920`: dataset of Financial Aid Statistics for 2019-2020
-   `staff_cat`: dataset of Staff Categories based on `admin2020$staff_cat`

## Example: Admissions and Test Scores for Fall 2020

```{r}
head(adm2020[,c(1:2, 11, 17, 30, 31, 34, 35)], 5)
```

## Example: Directory Information for 2020

```{r}
head(dir_info2020[,c(1:2,4:5)],5) 
```

## Functionality

```{r, results='hide'}
admission_reqs # extracts admission requirements for any institution
school_preferences # provides a dataset of institution according to your preferences
```

```{r, }
admission_reqs("Smith College")
```

```{r, include=FALSE}
library(dplyr)
```

```{r}
school_preferences(financial_aid = 70, size = 2, region = "New England", diversity_staff = 27) %>% select(c(1:3),6)
```
For further examples and capabilities, please see the [vignette]

## Contributors

- [Aushanae Haller](https://github.com/ajhaller)
- [Alejandra Munoz Garcia](https://github.com/alejanmg)
