---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# IPEDS

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/IPEDS)](https://CRAN.R-project.org/package=IPEDS)
<!-- badges: end -->

`IPEDS` is a package that contains datasets from The Integrated Postsecondary Education Data System (IPEDS). IPEDS is a system of interrelated surveys conducted annually by the U.S. Department of Education's National Center for Education Statistics (NCES). IPEDS annually gathers information from about 6,400 colleges, universities, and technical and vocational institutions that participate in the federal student aid programs.

## Installation

You can install the development version of IPEDS from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ajhaller/IPEDS")
```

## Load

```{r}
library(IPEDS)
```

## Available Tables

`adm2020`: dataset of Admissions and Test Scores for Fall 2020

## Example: Admissions and Test Scores for Fall 2020

```{r}
head(adm2020[,c(1:2, 11, 17, 30, 31, 34, 35)], 5)
```
## Contributors

- [Aushanae Haller](https://github.com/ajhaller)
- [Alejandra Munoz Garcia](https://github.com/alejanmg)
